Perturbing SU($n$)-symmetric interactions

### 2019-10-30 collective_sun/sun_perturbation.tex
generalized the two-body perturbation to the case of different operators on the addressed spins

We consider an array of $N$ multilevel spins with non-uniform
SU($n$)-symmetric interactions that can be written in the form
\begin{align}
  H_0 = \sum_{p<q} h_{pq} \Pi_{pq},
  &&
  \Pi_{pq} \equiv \sum_{\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)},
\end{align}
where $S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin
$p$ to $\ket\mu$ from $\ket\nu$; the operator $\Pi_{pq}$ permutes
spins $p$ and $q$; and the coefficients $h_{pq}$ satisfy
\begin{align}
  h_{pp} = 0
  &&
  h_{pq} \le 0,
  &&
  h_{pq} = h_{qp}.
\end{align}
The ground-state manifold of the interaction Hamiltonian $H_0$
consists of fully symmetric states that are simultaneous $+1$
eigenstates of all permutation operators $\Pi_{pq}$.  The energy of
these states is
\begin{align}
  E_0 \equiv \sum_{p<q} h_{pq} = \f12 \sum_{p,q} h_{pq}
  = \f12 \sum_p h_p = \f12 N h,
\end{align}
where we define
\begin{align}
  h_p \equiv \sum_q h_{pq},
  &&
  h \equiv \EE_p\sp{h_p} = \f1N \sum_{p,q} h_{pq},
  &&
  \EE_p\sp{X_p} \equiv \f1N \sum_p X_p.
\end{align}
We wish to determine the effective dynamics induced on the manifold
$\M_0$ of fully symmetric states by weak perturbations of the form
\begin{align}
  \V_1 \equiv \sum_p v_p V_p,
  &&
  \V_2 \equiv \sum_{p<q} v_{pq} V_p V_q,
  \label{eq:perturbations}
\end{align}
where $V_p$ is a single-body operator on spin $p$, and
$v_{pq}=v_{qp}$.  The effective Hamiltonian $H_W$ induced on the
ground-state manifold $\M_0$ by a perturbation $\V_W$ through second
order in perturbation theory is given by\cite{bravyi2011schrieffer,
  perlin2019effective}
\begin{align}
  H_W = H_W^{(1)} + H_W^{(2)},
  &&
  H_W^{(1)} = \P_0 \V_W \P_0,
  &&
  H_W^{(2)} = - \P_0 \V_W \E \V_W \P_0,
  &&
  \E \equiv \sum_{k>0} \f{\P_k}{\Delta_k},
\end{align}
where $\P_k$ is a projector onto the $k$-th excited eigenspace of the
interaction Hamiltonian $H_0$, with interaction energy $\Delta_k$
above that of fully symmetric manifold $\M_0$.

@@

\eqref{eq:perturbations} is

@@

  = \sum_{p<q} v_{pq} \P_0 V_p V_q \P_0
  \simeq \f{v}{N-1} \p{V^2 - W},

@@

@@

where we made use of the previous result in \eqref{eq:PVVP_neq_ops}.
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand

@@

  = \sum_{\substack{k<\ell\\p<q}} v_{k\ell} h_{pq}
  \Pi_{pq} V_k V_\ell \ket\psi

@@

  \Pi_{pq} V_k V_\ell \ket\psi.

@@

  \p{2 h_{k\ell} - h_k - h_\ell} v_{k\ell} V_k V_\ell \ket\psi,

@@

  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  v_{k\ell} \p{h_{kq} V_q V_\ell + h_{\ell q} V_k V_q} \ket\psi
  = \sum_{k,\ell,q} v_{k\ell} h_{kq} V_q V_\ell \ket\psi
  - \sum_{k,\ell} v_{k\ell} h_{k\ell} V_k^2 \ket\psi.

@@

    + \sum_q h_{kq} v_{q\ell}} V_k V_\ell \ket\psi
  - \sum_{k,\ell} h_{k\ell} v_{k\ell} V_k^2 \ket\psi.

>>>>

We consider an array of $N$ multilevel spins with non-uniform
SU($n$)-symmetric interactions that can be written in the form
\begin{align}
  H_0 = \sum_{p<q} h_{pq} \Pi_{pq},
  &&
  \Pi_{pq} \equiv \sum_{\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)},
\end{align}
where $S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin
$p$ to $\ket\mu$ from $\ket\nu$; the operator $\Pi_{pq}$ permutes
spins $p$ and $q$; and the coefficients $h_{pq}$ satisfy
\begin{align}
  h_{pp} = 0
  &&
  h_{pq} \le 0,
  &&
  h_{pq} = h_{qp}.
\end{align}
The ground-state manifold of the interaction Hamiltonian $H_0$
consists of fully symmetric states that are simultaneous $+1$
eigenstates of all permutation operators $\Pi_{pq}$.  The energy of
these states is
\begin{align}
  E_0 \equiv \sum_{p<q} h_{pq} = \f12 \sum_{p,q} h_{pq}
  = \f12 \sum_p h_p = \f12 N h,
\end{align}
where we define
\begin{align}
  h_p \equiv \sum_q h_{pq},
  &&
  h \equiv \EE_p\sp{h_p} = \f1N \sum_{p,q} h_{pq},
  &&
  \EE_p\sp{X_p} \equiv \f1N \sum_p X_p.
\end{align}
We wish to determine the effective dynamics induced on the manifold
$\M_0$ of fully symmetric states by weak perturbations of the form
\begin{align}
  \V_1 \equiv \sum_p v_p V_p,
  &&
  \V_2 \equiv \f12 \sum_{p,q} v_{pq} V_p \tilde V_q,
  \label{eq:perturbations}
\end{align}
where $V_p$ is a single-body operator on spin $p$, and we enforce
$v_{pq}=v_{qp}$ with $v_{pp}=0$.  The effective Hamiltonian $H_B$
induced on the ground-state manifold $\M_0$ by a perturbation $\V_B$
through second order in perturbation theory is given
by\cite{bravyi2011schrieffer, perlin2019effective}
\begin{align}
  H_B = H_B^{(1)} + H_B^{(2)},
  &&
  H_B^{(1)} = \P_0 \V_B \P_0,
  &&
  H_B^{(2)} = - \P_0 \V_B \E \V_B \P_0,
  &&
  \E \equiv \sum_{k>0} \f{\P_k}{\Delta_k},
\end{align}
where $\P_k$ is a projector onto the $k$-th excited eigenspace of the
interaction Hamiltonian $H_0$, with interaction energy $\Delta_k$
above that of fully symmetric manifold $\M_0$.

@@

\eqref{eq:perturbations} can be computed using the same tricks as for
$H_1^{(2)}$:

@@

  = \f12 \sum_{p,q} v_{pq} \P_0 V_p \tilde V_q \P_0
  \simeq \f{v}{N-1} \p{\f12 \sp{V,\tilde V}_+ - Z},
where $V,\tilde V,Z$ are respectively collective operators that sum
over the index of $V_p,\tilde V_p,Z_p$, and

@@

  Z_p \equiv \f12 \sp{V_p, \tilde V_p}_+,

@@

with $\sp{X,Y}_+\equiv XY+YX$ an anti-commutator.  In order to compute
the second-order effective Hamiltonian $H_2^{(2)}$, as before we pick
an arbitrary fully symmetric state $\ket\psi\in\M_0$ and expand

@@

@@

  \Pi_{pq} V_k \tilde V_\ell \ket\psi.

@@

  \p{2 h_{k\ell} - h_k - h_\ell}
  v_{k\ell} V_k \tilde V_\ell \ket\psi,

@@

  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} v_{k\ell}
  \p{h_{kq} V_q \tilde V_\ell + h_{\ell q} V_k \tilde V_q} \ket\psi
  = \sum_{k,\ell,q} v_{k\ell} h_{kq} V_q \tilde V_\ell \ket\psi
  - \sum_{k,\ell} v_{k\ell} h_{k\ell} Z_k \ket\psi.

@@

    + \sum_q h_{kq} v_{q\ell}} V_k \tilde V_\ell \ket\psi
  - \sum_{k,\ell} h_{k\ell} v_{k\ell} V_k \tilde V_k \ket\psi.
